\chapter{Model Design}\label{ch:modeldesign}
In this chapter, I discuss my research efforts toward developing a
model of the pthread library in SMACK.  I begin by introducing the
challenge of modeling the execution environment presented by the
operating system to the pthread library.  Having introduced this, I
proceed to discuss the process of modeling concurrency in the pthread
library over the lifecycle of a pthread.

\section{Execution Environment}
One of the critical tasks in implementing support for pthreads in
SMACK was modeling the environment that the operating system provides.
For each thread, the OS tracks the thread ID and thread status of
each thread spawned within a process.  This information is tracked by
the operating system in a thread control block (TCB).  At any time, a
thread must be able to query for its thread ID.  Similarly, a 
parent thread must be able to query the operating system for the
running state threads it has spawned.

Faithfully modeling the environment seen by the pthread library
requires addressing a number of bookkeeping issues.  These include
tracking the running state of each thread, and coordinating the thread
ID between the C level code and Corral itself.  The critical issue can
be illustrated by considering the API call
\lstinline|pthread_join(pthread_t* thr)|. This call blocks the calling
thread until the thread specified by \lstinline|thr| is in the stopped
state.  Consequently, properly model this function requires that the
calling thread must be able to query the environment for the status of
the thread referenced by \lstinline|thr|, in order to allow the
calling thread to unblock execution. 

Within the context of static model checking, because no code is
actually executing, the running state of a thread is simply a bit of
program state information -- a thread can be marked as waiting,
running, or stopped. Thus, tracking thread states requires the ability
to reference some bit of state information about the threads. A
\lstinline|pthread_t| object passed into \lstinline|pthread_join()|,
for example, must have some method of referencing this bit of thread
state information.  It seems natural to store each thread's thread ID
in this \lstinline|pthread_t| object, and use this to look up the
thread's state in an array.  Indeed, this is precisely how the
operating system tracks such information.  In addition, such an array
must be globally accessible, because threads must be able to query the
state of another thread (to singal \lstinline|pthread_join()|, for
example).  To model this, C level code is used to inject a global
Boogie-level array called \lstinline|$pthreadStatus|; the C level call

\begin{lstlisting}[frame=none,xleftmargin=2\parindent]
__SMACK_top_decl("var $pthreadStatus: [int][int]int;");
\end{lstlisting}

\noindent becomes

\begin{lstlisting}[frame=none,xleftmargin=2\parindent,language=boogie]
var $pthreadStatus: [int][int]int;
\end{lstlisting}

\noindent in the output Boogie program.

In order to access thread state from \lstinline|$pthreadStatus|, a
queried thread's thread ID is used to index into the array.  This
requires that a thread's ID is referenced by its
\lstinline|pthread_t| object.  This requires immediately querying
Corral for the thread ID of newly spawned thread, and recording this
information in its \lstinline|pthread_t|.  This is done from the
parent with \lstinline|corral_GetChildThreadID()|, and from the child
with \lstinline|corral_GetThreadID()|.  Now, when
\lstinline|pthread_join()|, for example, queries Corral for the
running state of a thread, it can pass in the thread ID to Corral,
which will use it to index into \lstinline|$pthreadStatus|.

However, there is further complication with implementing this
array. In Boogie, undefined variables take on a nondeterministic value
- that is, Corral assumes they can have any value.  Without
initializing \lstinline|$pthreadStatus|, Corral must consider the case
that threads start in a \emph{stopped} state.  This means that a call
to \lstinline|pthread_join()| immediately following a call to
\lstinline|pthread_create()| could query for the status of the newly
spawned thread before the new thread has initially set its status to
waiting, and see it is in the stopped state, causing the parent thread
to unblock.  To address this, \lstinline|$pthreadStatus| is
initialized so that all indices indicate an \emph{uninitialized}
state; the C level call 

\begin{lstlisting}[frame=none,xleftmargin=2\parindent]
__SMACK_code("assume (forall i:int :: "
                             "$pthreadStatus[i][0] == "
                             "$pthread_uninitialized);"); 
\end{lstlisting}

\noindent becomes 

\begin{lstlisting}[frame=none,xleftmargin=2\parindent,language=boogie]
assume (forall i:int :: $pthreadStatus[i][0] == 
                        $pthread_uninitialized);
\end{lstlisting}

\noindent in the output Boogie program.  This code gets executed as
the first statement in \lstinline|main()|, to ensure that any thread's
running state is defined prior to its use. 

With these Boogie level additions to each pthread program's Boogie
translation, the operating system environment is sufficiently prepared
to enable the modeling of the actual pthread library behavior.


\section{Thread Creation}
With the pthread library's execution environment modeled, a sufficient
framework exists to begin modeling the behavior of the actual pthread
library function calls.  In the pthread library, the procedure for
spawning threads is \lstinline|pthread_create()|.  There are two
aspects of \lstinline|pthread_create()| that require modeling:
behavior in the calling thread, and behavior in the newly spawned
thread. 

In the calling thread, there are two main tasks.  As the first task,
the new thread must be spawned, and instructed to execute the
procedure indicated by the \lstinline|*__start_routine| function
pointer.  As the second, SMACK must query Corral for the thread ID of
the newly spawned child thread, and record this value in the
\lstinline|pthread_t| struct for future reference.  This results in
the model defined in the C code seen in Figure~\ref{fig:pthread_create}.

\begin{figure}[h]
\centering
\caption{Model: pthread\_create()}\label{fig:pthread_create}
\begin{lstlisting}
int pthread_create(pthread_t* __newthread,
                   pthread_attr_t* __attr,
                   void* (*__start_routine) (void *),
                   void* __arg)
{
  __SMACK_code("async call @(@, @, @);", __call_wrapper,
                                         __newthread,
                                         __start_routine,
                                         __arg);

  __SMACK_code("call @ := corral_getChildThreadID();",
               *__newthread);

  return 0;
}
\end{lstlisting}
\end{figure}

In the newly spawned thread, again there is bookkeeping to do, as well
as executing the specified \lstinline|__start_routine()|.  Notice in
Figure~\ref{fig:pthread_create} that the function
\lstinline|__call_wrapper()| is the procedure being executed
asynchronously, rather than \lstinline|__start_routine()|.  This
allows for the child thread to perform bookkeeping before and after
calling \lstinline|__start_routine()|.

As seen in Figure~\ref{fig:__call_wrapper} the newly spawned thread
queries Corral for its thread ID, and then makes a call to
\lstinline|assume()|, which causes Corral to only consider paths
through the program where \lstinline|*__newthread| already contains
the thread ID of the child thread -- effectively, the new thread
``waits'' until its parent has set the child thread's ID in
\lstinline|*__newthread|. Once this has occurred, both the parent and
the child have a reference to the newly spawned thread's ID, and can
use this to index into \lstinline|$pthreadStatus| to set and query the
child thread's running state information.  

Once the child and parent can both reference the correct
\lstinline|$pthreadStatus| element, the new thread can proceed to
executing the specified \lstinline|__start_routine()|.  The new thread
cycles through the \emph{waiting} state to the \emph{running} state,
and then makes a synchronous call to the
\lstinline|__start_routine()|.  After the
\lstinline|__start_routine()| returns, the thread's running state is
finally set to \emph{stopped}, which signals the end of the child
thread. 

\begin{figure}[h]
\centering
\caption{Model: \_\_call\_wrapper()}\label{fig:__call_wrapper}
\begin{lstlisting}
void __call_wrapper(pthread_t* __newthread,
                    void* (*__start_routine)(void *),
                    void* __arg)
{
  int ctid = __SMACK_nondet();
  __SMACK_code("call @ := corral_getThreadID();", ctid);
  assume(ctid == *__newthread);
  
  __SMACK_code("$pthreadStatus[@][0] := $pthread_waiting;",
               *__newthread);
  __SMACK_code("$pthreadStatus[@][0] := $pthread_running;",
               *__newthread);

  __start_routine(__arg);

  __SMACK_code("$pthreadStatus[@][0] := $pthread_stopped;",
               *__newthread);
}
\end{lstlisting}
\end{figure}



\section{Thread Execution}
[Good]
With the ability to spawn threads, [wc it becomes time to] consider
the functionality of the pthread library that pertains to an executing
thread.  Among this functionality, the most commonly utilized are the
synchronization routines.

Corral extends Boogie IVL with two primitives for achieving
synchronization between threads: \lstinline|corral_atomic_begin()| and
\lstinline|corral_atomic_end()|.  These calls provide the ability to
create critical sections of Boogie IVL code, where only one atomic
block can execute at any time.  However, these atomic blocks are
global, and take no parameters.  There is no way to differentiate
atomic blocks of unrelated code that should be allowed to run
concurrently -- this is much like the big kernel lock in Linux.  This
presents the need to model the more complex synchronization routines
of pthreads using Corral's atomic primitives.

\subsection{Modeling Mutexes}
[Good]
Mutexes are the most commonly used mechanisms for providing
synchronization among threads.  In addition, they can be used as
building blocks to devise more complex synchronization mechanisms.
Modeling mutexes in Corral is relatively straightforward, and more
complex synchronization functionality easily builds on this
foundation. 

Each \lstinline|pthread_mutex_t| object allocated creates an
independent mutual exclusion device.  To track the current owner of a
mutex and prevent execution of threads that don't hold the lock, the
model stores the thread ID of the current mutex owner in the
\lstinline|pthread_mutex_t| struct.  This allows
\lstinline|pthread_mutex_lock()| to block until until the mutex is in
the \emph{UNLOCKED} state.


\begin{figure}[h]
\centering
\caption{Model: pthread\_mutex\_lock()}\label{fig:pthread_mutex_lock}
\begin{lstlisting}
int pthread_mutex_lock(pthread_mutex_t* __mutex)
{
  int tid = (int)pthread_self();
  assert(__mutex->init == INITIALIZED);
  assert(__mutex->lock != tid);

  __SMACK_code("call corral_atomic_begin();");
  assume(__mutex->lock == UNLOCKED);
  __mutex->lock = tid;
  __SMACK_code("call corral_atomic_end();");

  return 0;
}
\end{lstlisting}
\end{figure}

This leads to the straightforward model shown in
Figure~\ref{fig:pthread_mutex_lock}. The calling thread queries for
its thread ID using \lstinline|pthread_self()| (which simply queries
Corral with a call to \lstinline|corral_GetThreadID()|, and returns the
result). The mutex is then checked to ensure it is in the initial
state, and that the calling thread is not already the owner of the
mutex.  With these criteria ensured, the Boogie translation atomically
waits for the mutex to enter the \emph{UNLOCKED} state and updates the
mutex owner to be the calling thread's ID.

\lstinline|pthread_mutex_unlock()|, seen in
Figure~\ref{fig:pthread_mutex_unlock}, releases a mutex, and is
essentially the inverse of \lstinline|pthread_mutex_lock()|; the model
ensures the mutex is initialized and the calling thread is the current
owner, and atomically sets the mutex as being unlocked. 

\begin{figure}[h]
\centering
\caption{Model: pthread\_mutex\_unlock()}
\label{fig:pthread_mutex_unlock}
\begin{lstlisting}
int pthread_mutex_unlock(pthread_mutex_t* __mutex)
{
  int tid = (int)pthread_self();
  assert(__mutex->init == INITIALIZED);
  assert(__mutex->lock == tid);

  __SMACK_code("call corral_atomic_begin();");
  __mutex->lock = UNLOCKED;
  __SMACK_code("call corral_atomic_end();");

  return 0;
}
\end{lstlisting}
\end{figure}

\subsection{Complex Synchronization}
[Short, but sufficient?]
The pthread library also includes a number of more complex
synchronization mechanisms.  These allow for more efficient runtime
behavior, in addition to providing commonly utilized synchronization
mechanisms.

\lstinline|pthread_cond_wait(pthread_cond_t* cond, |
\lstinline|pthread_mutex_t* mutex)|  is one such function that
provides more efficient runtime behavior.  It blocks the calling
thread until both the \lstinline|cond| predicate is true \emph{and}
\lstinline|mutex| is \emph{UNLOCKED}, at which point the thread is
signaled to resume. This avoids having to continually acquire the
mutex and evaluate the predicate.  \lstinline|pthread_barrier_wait()|
is a good example of a procedure providing expected functionality.  It
causes calling threads to block until all threads calling
\lstinline|pthread_barrier_wait()| have reached this instruction.
These more complex synchronization routines are modeled entirely as C
level code simply by using \lstinline|pthread_mutex_t|'s as building
blocks. 

\section{Thread Termination}
[Good]
As a thread's execution comes to an end, there is a bit of bookkeeping
to be done to wrap up the thread's lifecycle.  It is primarily this
termination handling that required the upfront work of modeling the
operating system environment prior to implementing support for thread
creation.

There are two main considerations for handling the termination of a
thread.  The first consideration is the passing of a return value from
a terminating thread.  The second is blocking execution of a waiting
thread, and retrieving the return value of the terminating thread
being waited on by the blocking thread.

\subsection{Modeling Thread Exit and Join}
[Could use another pass or two]
It is often necessary to pass a return value from a terminating thread
to some other thread.  For example, with a scatter/gather design
pattern, a parent thread will send data to multiple worker threads,
which return a computational result back to the parent.

The challenge with achieving this was that this return data has to
exist somewhere between the time that the terminating thread exits,
and the joining thread retrieves it.  This functionality was
implemented by extending the definition of the
\lstinline|$pthreadStatus| array to store a tuple, rather than a 
single value.  With this, the model can store a return value passed
into \lstinline|pthread_exit(void* value)| as the second tuple
element, in addition to storing thread running state as the first.

\begin{figure}[h]
\centering
\caption{Model: pthread\_exit()}\label{fig:pthread_exit}
\begin{lstlisting}
void pthread_exit(void *retval)
{
  pthread_t tid = __SMACK_nondet();
  tid = pthread_self();

  __SMACK_code("assert $pthreadStatus[@][0] == "
               "$pthread_running;",
               tid);

  __SMACK_code("$pthreadStatus[@][1] := @;", tid, retval);

  __SMACK_code("$pthreadStatus[@][0] := $pthread_stopped;",
               tid);
}
\end{lstlisting}
\end{figure}

The model of \lstinline|pthread_exit()|, seen in
Figure~\ref{fig:pthread_exit}, gets the calling thread's ID and
ensures the thread is still in a running state, saves the return value
pointer in \lstinline|$pthreadStatus|, and finally sets the running to
stopped.  


With the return value from threads now stored in
\lstinline|$pthreadStatus|, a method for accessing this value is
needed.  This is done through 
\lstinline|pthread_join(pthread_t* thr)|, which blocks the calling
thread until the thread referenced  by \lstinline|thr| terminates.
The value returned from \lstinline|pthread_join()| is the pointer
passed in to \lstinline|pthread_exit()| by the terminating thread.

The model designed for \lstinline|pthread_join()|, shown in
Figure~\ref{fig:pthread_join} implements this behavior.  A call to
\lstinline|assume()| causes the thread to block until the target
thread's running state is \emph{stopped}.  Once the terminating thread
is stopped, the waiting thread fetches the return value pointer from
the \lstinline|$pthreadStatus| array, which is then returned by the
call to \lstinline|pthread_join()|.

\begin{figure}[h]
\centering
\caption{Model: pthread\_join()}\label{fig:pthread_join}
\begin{lstlisting}
int pthread_join(pthread_t __th, void **__thread_return)
{
  __SMACK_code("assume $pthreadStatus[@][0] == "
               "$pthread_stopped;",
               __th);

  __SMACK_code("@ := $pthreadStatus[@][1];",
               *__thread_return, __th);

  return 0;
}

\end{lstlisting}
\end{figure}


[leftover from intro]


[Remaining is leftover stuff - may be useful when revising intro]
This is supported by the accuracy seen in the empirical results of
initial implementation testing.  My proposed research will extend
SMACK with a model of a common subset of the pthreads API, enabling
verification of C/C++ programs utilizing the Pthreads library.
Extending SMACK to provide support for programs using pthreads has not
only provided the SMACK community with support for a common
concurrency paradigm, but also presents an opportunity to investigate
general constructs useful for modeling concurrency.  These constructs,
which will necessarily result from extending SMACK, will lend
themselves to modeling other concurrency libraries like OpenMP and
MPI.  Further, the resulting constructs should be general enough for
use within other verification tools. 

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "thesis"
%%% End: 
