\chapter{Related Work}\label{thesis_relatedwork}
Though I am unaware of any publications specifically on modeling
pthreads for static analysis, there are several verification projects
that provide support for the pthread library.  These project include
both dynamic verification tools, as well as static analysis tools.

Insepct is one such dynamic verification tool that provides support
for verifying programs utilizing the pthread library.  It uses modern 
algorithms to reduce the search space of thread scheduling, and
dynamically executes this reduced set of thread schedules, looking for
bugs such as deadlocks and race conditions.  It works by instrumenting
the source code to communicate with a scheduler which explores each of
the reduced set of thread schedules~\cite{inspect}.

CIVL is an example of a static analysis tool that provides support for
pthreads.  CIVL, itself, is an extension of the C programming
language that provides primitives for modeling
concurrency~\cite{civl}.  Included in the CIVL project is a symbolic
execution tool that reasons about the models designed using CIVL
programming language.  As a static analysis tool that models the
behavior of the pthread library, it is much more similar to SMACK than
Inspect. 

Despite the existence of verification projects supporting programs
utilizing the pthread library, the extension to SMACK providing
support for pthread programs is justified.  SMACK is an actively
developed tool with an active and growing user base.  Extending SMACK
to support pthreads provides this community with a tool for analyzing
concurrent programs -- a feature previously lacking in SMACK.

Another promising project is a lock set analysis module for SMACK.
Under development by Pantazis Deligiannis, the tool tracks critical
sections locked with a mutex or spinlock, and prevents SMACK from
considering thread interleavings of mutually excluded critical
sections that do not have access conflicts on global variables, or
race conditions. This currently unpublished work should be easy to
integrate with the pthread support implemented for SMACK, and promises
to drastically reduce the number of thread interleavings that SMACK is
required to explore.  

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "thesis"
%%% End: 
